### 2020-07-16 Manuscript/draft.Rmd

---
title: "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(dplyr)
library(magrittr)
library(tidyr)
library(stringr)
library(citr)
library(ggplot2)
library(papaja)

```

# Abstract

Many studies in bilingualism have explored how form similarity between translation equivalents (i.e. cognateness) impacts language processing. Cognates range from being identical to sharing only a few key features. Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity. In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value. A widely used metric for word similarity is Levenshtein distance [@levenshtein1966binary]. This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other. However, a challenge when defining phonological similarity is that some phonemes (e.g. [k] and [g]) are perceptually more similar than others (e.g. [k] and [o]). The standard Levenshtein distance calculation treats each change equally. Thus it often overestimates phonological distance between words. We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g. vowel or consonant insertion) and position in the word. To estimate the weights, we are collecting behavioural data in an auditory translation elicitation task. Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words. The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent. Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding translation. The higher the probability, the higher the perceptual similarity. Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.

# Introduction

# Methods

## Computing cognateness: Phonological distance across translation equivalents

## Behavioural task: Translation elicitation task

### Sample

We collected data from XXX monolingual participants with no history of hearing, motor, or language impairments. XXX were English monolinguals (XXX females, age range = XXX, mean age = XXX), and XXX were Spanish monolinguals (XXX females, age range = XXX, mean age = XXX). English-monolingual participants were recruited in UK and Spanish monolinual particpiants were recruited in Spain. Participants were contacted via email, provided informed consent prior to any data collection, and were retributed with 7€/6X after completing the experiment. This experiment was approved by the Ethical Committee of XXXXXXX, with code XXXXXXXXXXXXXX.

The language profile of participants was assesed using a brief *ad hoc* questionnaire were participants reported previous experience and with any second language, if any. Participants were asked to report their proficiency in oral comprehension and orthography in Spanish, Catalan, English, and any other non-native languge, if reported. Participants with exposure to a non-native language for longer than XXX or a proficiency in any non-native language higher than 2 in a 5-point lickert scale were discarded.

### Stimuli

### Design

### Procedure

### Apparatus

# Results

# Discussion

# Appendix

## Appendix 1: Session info

```{r sessionInfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

sessionInfo()

```