### 2020-07-30 Manuscript/te_appendix.Rmd

---
title: "Appendices"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare}

prior <- fread(here("Results", "prior_predictions.csv"), na.strings = "") %>% as_tibble()
posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()

```

# Appendix 2: Model

Prior predictive checks

Convergence

```{r convergence}

posterior %>%
    filter(.iteration < 2000) %>% 
    ggplot(aes(.iteration, .value, colour = .chain)) +
    facet_wrap(~.variable, scales = "free_y") +
    geom_line() +
    labs(x = "Iteration", y = "Value", colour = "Chain") +
    theme_custom +
    theme(legend.position = "top") +
    ggsave(here("Figures", "convergence.png"))

```

# Appendix 4: Session info

```{r session_info}

sessionInfo()

```