### 2020-08-06 Communications/2020-07-30_meeting.Rmd
Prepare report for meeting

---
title: "Translation Elicitation"
date: "30/07/2020"
output: html_document
---

```{r prepare, include=FALSE}

library(dplyr)
library(papaja)
library(tidyr)
library(tibble)
library(janitor)
library(stringr)
library(forcats)
library(ggplot2)
library(patchwork)
library(here)
library(data.table)
library(broom)
library(brms)
library(effectsize)
library(pairwiseComparisons)
library(tidybayes)
library(gt)
library(lubridate)

participants <- fread(here("Data", "02_participants.csv"), na.strings = c("", "NA")) %>%
  as_tibble() %>%
  mutate(date = as_date(date),
         group = as.factor(group))
contrasts(participants$group) <-  contr.helmert(3)/2

trials <- fread(here("Data", "trial_stats.csv"),
                na.strings = c("", "NA")) %>%
  as_tibble() %>%
  clean_names()

accuracy <- fread(here("Data", "03_accuracy_coded.csv"),
                  na.strings = c("", "NA")) %>%
  as_tibble() %>% 
  mutate(correct_coded = response_type %in% c("correct", "typo"),
         correct_coded = as.numeric(correct_coded))

posterior <- fread(here("Results", "posterior_draws.csv"),
                   na.strings = c("", "NA")) %>%
  as_tibble()

prior <- fread(here("Results", "prior_draws.csv"),
               na.strings = c("", "NA")) %>%
  as_tibble()

predictions <- fread(here("Results", "posterior_predictions.csv"),
                     na.strings = c("", "NA")) %>%
  as_tibble()

fit <- readRDS(here("Results", "fit4_phon.rds"))
loos <- readRDS(here("Results", "loos.rds"))
r2 <- readRDS(here("Results", "r2.rds"))

item_effects <- fread(here("Results", "item_effects.csv"),
                      na.strings = c("", "NA")) %>%
  as_tibble()

group_effects <- fread(here("Results", "group_effects.csv"),
                       na.strings = c("", "NA")) %>%
  as_tibble()

corrs <- fread(here("Results", "corrs.csv"),
               na.strings = c("", "NA")) %>%
  as_tibble()

valid_participant <- filter(participants, valid_participant) %>% pull(participant)

clean_varnames <- function(x){
  str_remove({{ x }}, "b_") %>% 
    str_replace_all(., "pthn", "PTHN") %>% 
    str_replace_all(., "similarity_phon", "Phon. Similarity") %>% 
    str_replace_all(., ":", " \U000D7 ")
}

```

# Behavioural task

## Sample

`r nrow(participants)` participants (mean age = `r printnum(mean(participants$age, na.rm = TRUE))`, *SD* age = `r printnum(sd(participants$age, na.rm = TRUE))`, age range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).

`r filter(participants, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(participants, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(participants, country=="Spain") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(participants, country=="Spain", sex =="Female") %>% nrow()` female).

Exclusion criteria:

a) Oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`)
b) Language impairment (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION? (Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan? (Sí/No)". However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%). This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.]
c) Did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`)

This resulted in `r nrow(participants)-sum(participants$valid_participant)` participants excluded, leaving `r nrow(filter(participants, valid_participant))` valid participants and `r accuracy %>% filter(response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% nrow() %>% printnum(big.mark = ",")` valid trials.

```{r participants_prof_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>%
  filter(!(l2 %in% "None")) %>% 
  pivot_longer(c(l2oral, l2written), names_to = "modality", values_to = "proficiency") %>%
  mutate(modality = str_to_sentence(str_remove(modality, "l2"))) %>%
  group_by(l2, country, modality) %>% 
  summarise(mean = mean(proficiency, na.rm = TRUE),
            n = n(), .groups = "drop") %>%
  arrange(n) %>% 
  mutate(l2 = fct_inorder(as.factor(l2))) %>% 
  ggplot(aes(l2, mean, fill = l2)) +
  facet_grid(country~modality) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(y = mean-0.5, label = paste0("n = ", n)), size = 3) +
  labs(x = "L2", y = "Self-rated proficiency (0-5)") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(limits = c(0, 5)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        axis.title.x = element_blank(),
        axis.text = element_text(colour = "black"))

```

```{r participants_prof_spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>%
  pivot_longer(c(spanish_oral, spanish_written, catalan_oral, catalan_written), names_to = "language", values_to = "proficiency") %>%
  separate(language, c("language", "modality"), sep = "_") %>% 
  mutate(modality = str_to_sentence(str_remove(modality, "l2")),
         language = as.factor(str_to_sentence(language))) %>% 
  group_by(language, country, modality, proficiency) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  ggplot(aes(x = proficiency, y = n, fill = country)) +
  facet_grid(language~modality) +
  geom_col() +
  labs(y = "Count", x = "Self rated proficiency (0-5)", fill = "Country") +
  scale_x_continuous(limits = c(0, 5)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        axis.text = element_text(colour = "black"))

```

## Design

* 0 - 1000 ms: fixation point
* 1000 - 2500 ms: audio
* 2500 ms - Press RETURN key

## Stimuli

```{r stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

plot_onset <- trials %>%
  select(trial_id, word, language, matches("same_onset")) %>% 
  rename_all(str_remove, "same_") %>% 
  pivot_longer(matches("onset"), names_to = "language_pair", values_to = "onset") %>%  
  mutate(language_pair = str_to_upper(str_remove(language_pair, "onset_")),
         onset = ifelse(str_detect(onset, "same"), "Same", "Different")) %>% 
  drop_na(onset) %>% 
  ggplot(aes(onset, colour = language_pair, fill = language_pair)) +
  facet_wrap(~language_pair) +
  geom_bar(position = position_dodge(width = 1), na.rm = TRUE) +
  labs(x = "Onset", y = "Counts",
       colour = "Language pair", fill = "Language pair") +
  theme_classic() +
  theme(legend.position = "none")

plot_subs <- trials %>%
  select(trial_id, word, language, matches("close_")) %>% 
  rename_all(str_remove, "close_") %>% 
  pivot_longer(matches("substitutions"), names_to = "language_pair", values_to = "substitutions") %>% 
  mutate(language_pair = str_to_upper(str_remove(language_pair, "substitutions_"))) %>% 
  ggplot(aes(language_pair, substitutions, colour = language_pair, fill = language_pair)) +
  geom_violin() +
  stat_summary(fun = "mean", geom = "point", colour = "black", shape = "_", size = 10) +
  labs(x = "Close substitutions", y = "Phonological similarity",
       colour = "Language pair", fill = "Language pair") +
  scale_y_continuous(limits = c(0, 1)) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.x = element_blank())

plot_sim <- trials %>%
  select(trial_id, word, language, matches("similarity")) %>% 
  rename_all(str_remove, "similarity_") %>% 
  rename_at(vars(4:9), ~paste0("phon_", .)) %>% 
  rename_at(vars(4:9), str_replace, "phon_ort_", "ort_") %>% 
  pivot_longer(4:9, names_to = c("modality", "language_pair"), values_to = "similarity", names_sep = "_") %>% 
  pivot_wider(names_from = modality, values_from = similarity) %>% 
  mutate(language_pair = str_to_upper(language_pair)) %>% 
  ggplot(aes(ort, phon, colour = language_pair, fill = language_pair)) +
  geom_point(alpha = 0.5, size = 0.25) +
  geom_smooth(formula = "y ~ x", method = "lm") +
  labs(x = "Orthographical similarity", y = "Phonological similarity",
       colour = "Language pair", fill = "Language pair") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  theme_classic()

plot_onset + plot_subs + plot_sim + guide_area() +
  plot_layout(nrow = 2, guides = "collect") &
  theme(axis.text = element_text(colour = "black"),
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))

```
## Data analysis

We fit Bayesian logistic models to model the probability of participants typing a correct translation in each trial. As predictors, we included the number of more frequent phonological neighbours of the translation (`PTHN`, standardised), whether both translation equivalents started with the same phoneme (`same_onset`, with levels `different`, `same_vowel`, `same_consonant`, sum-coded as `different` (-0.5) vs. `same` (0.5), the highest number of consecutive phonemes shared (`consecutive_longest`, standardised), and our measure of phonological similarity (`similarity_phop`, standardised). We first for a null model `fit0_null` that only included PTHN as predictor. We then extended the model incrementally by adding other predictors. The models were finally, we fitted a model that only included `PTHN` and the traditional orthographical Levenshtein similarity (`similarity`, standardised), we test whether our measure predicted participants responses better than just relying on orthographic similarity. The models were implemented in the `brms` R package using the following formulas:

* `fit1_pthn <- correct | trials(total) ~ pthn + (1 + pthn | group)`
* `fit2_onset <- correct | trials(total) ~ pthn*same_onset + (1 +  pthn*same_onset | group)`
* `fit3_close <- correct | trials(total) ~ pthn*close_substitutions + (1 + pthn*close_substitutions | group)`
* `fit4_phon <- correct | trials(total) ~ pthn*similarity_phon + (1 + pthn*similarity_phon | group)`
* `fit5_ort <- correct | trials(total) ~ pthn*similarity_ort + (1 + pthn*similarity_ort | group)`

# Results

```{r accuracy_anova, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

model <- lm(prop_correct ~ group, data = participants)
anov <- anova(model) %>% 
  tidy()

es <- anov$sumsq[1]/(anov$sumsq[1]+anov$sumsq[2])
pairs <- pairwise_comparisons(data = participants,
                              x = group,
                              y = prop_correct, type = "parametric",
                              p.adjust.method = "bonferroni") %>% 
  select(-c(label, test.details, p.value.adjustment, significance))

```

We performed a one-way ANOVA to analyse how participants' accuracy (the proportion of correct trials) changed across groups the three groups (ENG-SPA, ENG-CAT, and SPA-CAT). The `group` predictor explained a significant proportion of the variance, *F*(`r printnum(anov$df[1])`, `r printnum(anov$df[2])`) = `r printnum(anov$statistic[1])`, `r printp(anov$p.value[1])`, $\hat{\eta}^2$ = `r printnum(es)`). Follow-up *t*-tests revealed that Spanish-native speakers translating Catalan words were more accurate than English-native participants translating Spanish words, *t*(`r printnum(pairs$df[3])`) = `r printnum(pairs$t.value[3])`, `r printp(pairs$p.value[3])`, Cohen's *d* = `r printnum(pairs$t.value[3]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`), or Catalan words, *t*(`r printnum(pairs$df[2])`) = `r printnum(pairs$t.value[2])`, `r printp(pairs$p.value[2])`, Cohen's *d* = `r printnum(pairs$t.value[2]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`). We did not find evidence against accuracy being similar in both English-native groups, , *t*(`r printnum(pairs$df[1])`) = `r printnum(pairs$t.value[1])`, `r printp(pairs$p.value[1])`, Cohen's *d* = `r printnum(pairs$t.value[1]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`). Bonferroni corrections were applied in the aforementioned contrasts.

```{r accuracy, echo=FALSE, fig.cap="Participants' overall accuracy across groups. The X-axis represents the native language of participants (English or Spanish) and the language participants listened during the task (Spanish or Catalan). The Y-axis represents the proportion of correct. Each datapoint corresponds to one participant. Horizontal black lines represent group means."}

ggplot(participants, aes(group, prop_correct, fill = group)) +
  geom_violin(colour = NA) +
  geom_boxplot(fill = "white", width = 0.1) +
  labs(x = "Group", y = "Accuracy (% correct trials)", fill = "Group") +
  scale_y_continuous(limits = c(0, 1)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.text = element_text(colour = "black"))
```

## Model comparison

```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

r2_table <- r2 %>%
  mutate(r2_ci95 = paste0("[", round(q2_5, 2), ", ", round(q97_5, 2), "]")) %>% 
  rename(r2 = estimate, r2_se = est_error) %>% 
  select(model, r2, r2_se, r2_ci95)
loos$diffs %>% 
  as.data.frame() %>% 
  rownames_to_column("model") %>% 
  select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>% 
  left_join(r2_table) %>% 
  relocate(model, r2, r2_se, r2_ci95) %>% 
  gt() %>% 
  fmt_number(c(2, 3, 5, 6, 7, 8)) %>% 
  cols_label(model = "Model", r2 = "R2", r2_se = "SE", r2_ci95 = "95% CrI", elpd_diff = "\U0394 ELPD", se_diff = "\U0394 SE", elpd_loo = "ELPD LOO", se_elpd_loo = "SE ELPD LOO") %>%
  tab_spanner(label = "R2", columns = matches("r2")) %>% 
  tab_spanner(label = "PSIS-LOO", columns = matches("loo|diff"))

```

## Posterior distributions

```{r posterior_coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

summary(fit)$fixed %>%
  as.data.frame() %>% 
  rownames_to_column("term") %>% 
  clean_names() %>% 
  mutate(ci95 = paste0("[", round(l_95_percent_ci, 2), ", ", round(u_95_percent_ci, 2), "]"),
         term = clean_varnames(term)) %>% 
  select(-matches("percent")) %>% 
  relocate(term, estimate, est_error, ci95) %>% 
  gt(rowname_col = "term") %>% 
  fmt_number(columns = vars(estimate, est_error, rhat, bulk_ess, tail_ess)) %>% 
  cols_label(estimate = "Estimate", est_error = "SE", ci95 = "95% CrI", rhat = "R-hat", bulk_ess = "Bulk ESS", tail_ess = "Tail ESS")

```

```{r posterior_draws, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Posterior distribution of the extended model's (fit3_phon) coefficients. The X-axis represents all the values that each coefficient can take. The Y-axis represents the probability density at each value, i.e., the probability of the coefficient taking a specific value. Grey bars colour-code highest density intervals (HDI) of different width. Each interval indicates the narrowest interval that can be define within each distribution capturing a given proportion of the probability mass (e.g., given our data there is a 95% probability that the true value of a coefficient lies within the 95% HDI interval."}

posterior %>% 
  mutate(.variable = clean_varnames(.variable)) %>% 
  ggplot(aes(.value, .variable)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(show.legend = FALSE, fill = "grey30") + 
  stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.2)) +
  labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
       title = "Posterior distribution of coefficients of fixed effects") +
  scale_colour_brewer(palette = "Oranges") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"),
        legend.position = "top",
        legend.key = element_rect(fill = "transparent", colour = NA),
        axis.text = element_text(colour = "black"))

```

## Posterior predictions

```{r posterior_predictions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Predicted probability of correct translation across different measures of phonological similarity scores. Predictions were generated by sampling 1000 random draws from the posterior distribution of each predictor in the extended model (fit3_phon)."}

predictions %>%
  ggplot(aes(similarity_phon, .value)) +
  geom_point(data = accuracy, aes(y = correct_coded), shape = "|") +
  geom_line(aes(group = .draw), alpha = 0.25, colour = "grey") +
  stat_summary(fun = "mean", geom = "line", colour = "blue") +
  labs(x = "Phonological similarity", y = "Probability of correct translation",
       fill = "PTHN", colour = "Group") +
  theme_classic() +
  theme(axis.text = element_text(colour = "black"))
```

## Random effects

```{r group_effects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Group-level effects. We specified a maximal random effects structure. As a result, intercepts and slopes were estimated for each group. This figure shows how the estimated intercepts and slopes varied across groups. There are not any remarkable differences across groups, except for the slope of phonological similarity. The accuracy of Spanish-native speakers translating from Catala \\(SPA-CAT\\) was slightly boosted by the phonological similarity between translation equivalents in comparison with the rest of the groups \\(ENG-SPA, ENG-CAT\\)."}

group_effects %>%
  mutate(param = clean_varnames(param)) %>% 
  ggplot(aes(.value, group)) +
  facet_wrap(~param, scales = "free") +
  stat_interval(.width = c(0.95, 0.89, 0.67, 0.50)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Estimate", y = "Group", colour = "HDI (%)") +
  scale_colour_brewer() +
  theme_classic() +
  theme(legend.position = "top",
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.title.y = element_blank(),
        axis.text = element_text(colour = "black"))

```

```{r correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Pairwise correlations between group-level effects. This figure displaya"}

corrs %>% 
  mutate(.variable = str_remove(.variable, "cor_group__"),
         .variable = str_replace(.variable, "__" , " - "),
         .variable = clean_varnames(.variable)) %>% 
  ggplot(aes(.value, .variable)) +
  stat_interval(.width = c(0.95, 0.89, 0.67, 0.50)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Estimate", y = "Group", colour = "HDI (%)") +
  scale_colour_brewer(palette = "Oranges") +
  theme_classic() +
  theme(legend.position = "top",
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.title.y = element_blank(),
        axis.text = element_text(colour = "black"))

```