### 2021-05-20 Rmd/compare_jtrace_translation.Rmd
Integrate jtRACE script by @SiowSerene into the main report, clean repository, update data files.

---
title: "Comparing jTRACE Output with Translation Elicitation Task"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggplot2)

```

```{r load jtrace output}
filedir <- "data/" 
# directory with save files (
  ## 1 file per input word, with 1 output word per column
  ## files should be named with input word orthography and language (e.g. "gato_spanish.xlsx")
  ## files should be xlsx (csv doesn't deal well with special characters like ^ and @)
  ## folder should not contain any other files

# Note: it's effortful to have to convert to xlsx (jTRACE doesn't allow direct xlsx export), but it's the most reliable method to preserve special characters I've found so far. Would welcome ideas

# Export jTRACE output with Top 100 items and 100 cycles (arbritary value)
# Max ad-hoc activation

fnames <- dir(filedir) # extract list of file names in directory

trace_output <- data.frame() # create empty data frame

# for loop to extract highest activations for each input word
for (i in 1:length(fnames)){
  filename <- paste(filedir, fnames[i], sep = "") # get file name
  temp <- read_excel(filename) # read file

  temp <- apply(temp,2,max) # get highest value in each column
  temp <- as.data.frame(temp)
  temp <- cbind(newColName = rownames(temp), temp) # change index column (which has output items) into column 1 for easier organisation
  rownames(temp) <- 1:nrow(temp)
  names(temp)[1] <- "output" # (rename col) output item
  names(temp)[2] <- "activation" # (rename col) max activation value

  temp <- temp %>%
    filter(!output == "cycle") %>% # remove row corresponding to cycle number
    mutate(input = fnames[i]) %>% # specify input word
    arrange(desc(activation)) %>%
    filter(!activation <= 0) # remove low activation (maybe even apply a higher threshold? TBD)

  trace_output <- bind_rows(trace_output, temp)
}

trace_output <- trace_output %>%
  mutate(input = str_remove(input, ".xlsx")) %>%
  separate(input, c("input", "language"), "_")

```

```{r match word}

# xlsx file with 1 column for TRACE transcriptions and 1 column for orthography (English words) 
df_transcription <- read_excel("trace_transcription.xlsx") %>%
  select(trace_eng, ort_eng) %>%
  unique()

# Match the jTRACE output transcriptions with their orthography
trace_output <- left_join(trace_output, df_transcription, by = c("output" = "trace_eng")) %>%
  rename(output_word = ort_eng)

```

```{r load translation task}

# Cleaned file with participant answers, 1 row per trial per participant
TranslationTask <- read.csv("translation_engdata_cleaned.csv") %>%
  mutate(test_language = tolower(test_language))

# count the number of participants who gave each answer in response to a given trial
df_answer <- TranslationTask %>%
  group_by(test_language, word, answer) %>%
  summarise(n_participant = n(), .groups = "drop") %>%
  group_by(test_language, word) %>%
  mutate(pct_participant = n_participant/sum(n_participant)*100) #%>%
#  filter(!n_participant == 1) # remove answers only given by 1 participant as idiosyncratic answers?

```

### Predicting participant response by jTRACE

```{r plot, warning = FALSE}

# Predicting participant response by jTRACE

df_match <- full_join(trace_output, df_answer, by = c("input" = "word", "output_word" = "answer", "language" = "test_language")) %>%
  semi_join(., trace_output, by = c("input", "language"))

# answers
print("Presented words and the answer that was given by the most number of participants for that trial")

# If "output" is <NA>, the word is not in the currently-used jTRACE lexicon
df_match %>%
  filter(language == "spanish") %>% 
  arrange(input, desc(pct_participant)) %>%
  group_by(input) %>%
  slice(1) %>%
  select(input, output_word, pct_participant, output, activation) %>%
  arrange(desc(pct_participant))

df_match %>%
  filter(language == "catalan") %>% 
  arrange(input, desc(pct_participant)) %>%
  group_by(input) %>%
  slice(1) %>%
  select(input, output_word, pct_participant, output, activation) %>%
  arrange(desc(pct_participant))

# if no data for a given output word is available in df_answer, it means no participants gave that word as a response - recode as 0
df_match$pct_participant[is.na(df_match$pct_participant)] <- 0

# if no data for a given output word is available in trace output, it means that activation is outside of Top 100 words and/or below 0
df_match$activation[is.na(df_match$activation)] <- 0

# However, this is also constrained by the size of TRACE's lexicon, so subset only the words that do appear in the lexicon
df_match_trace <- semi_join(df_match, df_transcription, by = c("output_word" = "ort_eng"))

# plot
ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
  geom_smooth() +
  geom_point() +
  ggtitle("jTRACE activation against % participants who gave word as response")

# correlation test
cor.test(df_match_trace$activation, df_match_trace$pct_participant)

# plot by word
ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
  geom_smooth() +
  geom_point() +
  xlab("jTRACE activation value for the word") +
  ylab("% participants who responded with the word") +
  ggtitle("jTRACE activation against % participants who gave word as response") +
  facet_wrap(language~input)

```