### 2021-05-25 Rmd/jtrace_report.Rmd
Add model comparisons using LOO-IC

---
title: "jTRACE notbook"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = TRUE,
    warning = TRUE, 
    cache.extra = knitr::rand_seed,
    out.width = "100%",
    results = "asis",
    dev.args = list(png = list(type = "cairo"))
)
options(knitr.kable.NA = '-')
```

```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# load packages
library(tidyverse)
library(googlesheets4)
library(gt)

# import data
phonemes <- range_read(
    "1iAK2zF84MqFwdLj-_CpWVC1O2xxmYDztc0Qe_2Sy0ww",
    sheet = "(Gonzalo coding)"
) %>% 
    mutate_at(
        vars(pow, voc, diff, acu, cons, voi, burst),
        as.numeric
    )

lexicon <- sheet_names("1pZE58wHYGFxaIC90hSr-202Qv9go6-VZLixl_lF1QKY")[-1] %>% 
    map(
        function(x){
            range_read(
                ss = "1pZE58wHYGFxaIC90hSr-202Qv9go6-VZLixl_lF1QKY",
                sheet = x
            ) %>% 
                select(trial_id, ipa1, ipa2, word1, word2, jTRACE1, jTRACE2)
        }
    ) %>%
    set_names(sheet_names("1pZE58wHYGFxaIC90hSr-202Qv9go6-VZLixl_lF1QKY")[-1])

```

# Phonemes {.tabset .tab-pills}

## Phoneme inventory {.tabset .tab-pills}

### Consonants

```{r phonemes_consonants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Consonant") %>% 
    select(id, ipa, sampa, trace, english, spanish, catalan, voice, place, manner) %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(everything()) %>% 
    cols_label(
        id = "ID",
        sampa = "SAMPA",
        trace = "jTRACE",
        english = "English",
        spanish = "Spanish",
        catalan = "Catalan",
        voice = "Voice",
        place = "Place",
        manner = "Manner"
    ) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("english"),
			rows = english
		)
	) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("spanish"),
			rows = spanish
		)
	) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("catalan"),
			rows = catalan
		)
	) %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### Vowels

```{r phonemes_vowels, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Vowel") %>% 
    select(id, ipa, sampa, trace, english, spanish, catalan, height, backness, roundness) %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(2:7) %>% 
    cols_label(
        id = "ID",
        sampa = "SAMPA",
        trace = "jTRACE",
        height = "Height",
        backness = "Backness",
        roundness = "Roundness"
    ) %>% 
       tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("spanish"),
			rows = spanish
		)
	) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("catalan"),
			rows = catalan
		)
	) %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### {-}

## Features {.tabset .tab-pills}

### Consonants

```{r features_consonants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Consonant") %>% 
    select(id, ipa, trace, description, pow, cons, voc, diff, acu, voi, burst) %>% 
    gt() %>% 
    cols_label(
        id = "ID",
        ipa = "IPA",
        trace = "jTRACE",
        description = "Description",
        pow = "Power",
        voc = "Vocalic",
        diff = "Diffuseness",
        acu = "Acuteness",
        cons = "Consonantal",
        voi = "Voicing",
        burst = "Burst"
    ) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(2:4) %>%
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### Vowels

```{r features_vowels, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Vowel") %>% 
    select(id, ipa, trace, description, pow, cons, voc, diff, acu, voi, burst) %>% 
    gt() %>% 
    cols_label(
        id = "ID",
        ipa = "IPA",
        trace = "jTRACE",
        description = "Description",
        pow = "Power",
        voc = "Vocalic",
        diff = "Diffuseness",
        acu = "Acuteness",
        cons = "Consonantal",
        voi = "Voicing",
        burst = "Burst"
    ) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(2:4) %>%
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### Feature correlations

```{r features_correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    select(pow, voc, diff, acu, cons, voi, burst) %>% 
    mutate_all(replace_na, replace = 0) %>%
    as.matrix() %>% 
    heatmap(
        labRow = phonemes$trace,
        labCol = c("Power", "Vocalic", "Difusseness", "Acuteness", "Consonantal", "Voice", "Burst"),
        Colv = NA,
        na.rm = FALSE
    )
```

### {-}

## {-}

# Lexicon {.tabset .tab-pills}

## English-Spanish

```{r lexicon_engspa, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
lexicon$`English-Spanish` %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

## English-Catalan

```{r lexicon_engcat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
lexicon$`English-Catalan` %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

## Spanish-Catalan

```{r lexicon_spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
lexicon$`Spanish-Catalan` %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

## {-}

# Parameters

# Design