### 2021-07-01 Rmd/00_outline.Rmd
Rename files, include job, add index.html

---
title: "Outline"
date: "Updated: `r format(Sys.Date(), '%d/%m%/%Y')`"
output:
    pdf_document:
        latex_engine: xelatex
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The impact of cognateness in bilingual lexical acess

### Comprehension

### Production

### Translation

* [@dijkstra2010how]

### Phonological neighbourhoods and parallel activation

### Limitations of current measures of cognateness

* Levenshtein distance
* Coltheart N

## The present study