### 2021-09-29 Rmd/Old/02_methods.Rmd
Remove big files

---
title: "Methods"
date: "Updated: `r format(Sys.Date(), '%d/%m%/%Y')`"
output:
    pdf_document:
        latex_engine: xelatex
bibliography: "references.bib"
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

knitr::opts_chunk$set(
	out.width = "100%",
	message = FALSE,
	warning = FALSE
)

```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(papaja)
library(gt)
library(knitr)
library(lubridate)
library(here)

participants <- readRDS(here("Data", "participants.rds"))
stimuli <- readRDS(here("Data", "stimuli.rds"))
responses <- readRDS(here("Data", "responses.rds"))

```