### 2021-12-16 Rmd/appendix.Rmd
Fix error when joining stimuli info and participant responses

---

---

# Appendix 1: MCMC diagnostics

```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

suppressWarnings({
  bayesplot_theme_set(
    new = theme_custom() + 
      theme(
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8),
        strip.text = element_text(size = 10),
        legend.key = element_rect(colour = NA, fill = "white")
      )
  )
})

color_scheme_set("brightblue")
pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")
pars_labels <- c(
  "Intercept",
  "Frequency (+1 SD)",
  "PTHN \u00d7 Levenshtein", 
  "Levenshtein (+1 SD)",
  "PTHN (+1 SD)"
)
```

## Pairs plot

```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
posterior_cp <- as.array(model_fits$fit_4)
np_cp <- nuts_params(model_fits$fit_4)

plot_pairs <- mcmc_pairs(
  posterior_cp, pars = pars,
  off_diag_args = list(size = 0.75),
  diag_args = list(facet_args = list(labeller = label_value(labels = pars_labels)))
)

plot_pairs
```

## Trace and Rank plots

```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
(mcmc_trace(model_fits$fit_4, pars) + legend_none()) /
  mcmc_rank_overlay(model_fits$fit_4, pars = pars) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "top"
  )
```

## Gelman-Rubin convergence criterion (R-hat)

```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rhats <- rhat(model_fits$fit_4)
ratios <- neff_ratio(model_fits$fit_4)

mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
  mcmc_neff_hist(ratios, binwidth = 0.1) +
  plot_layout(guides = "keep", ncol = 1) &
  theme(
    legend.position = "top"
  )
```

# Appendix 3: participant-level posterior predictions

```{r acc_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# by participant
m_re_participant <- posterior_epreds_random$participants %>% 
  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
  mutate(
    pthn = case_when(
      pthn=="-1 SD" ~ "Low PTHN",
      pthn=="0 SD" ~ "Mean PTHN",
      pthn=="1 SD" ~ "High PTHN"
    ),
    frequency_zipf = case_when(
      frequency_zipf=="-1 SD" ~ "Low frequency",
      frequency_zipf=="0 SD" ~ "Mean frequency",
      frequency_zipf=="1 SD" ~ "High frequency"
    )
  ) %>% 
  group_by(participant, lv, pthn) %>% 
  summarise(.epred = mean(.epred, na.rm = TRUE), .groups = "drop") %>% 
  mutate(participant = as.numeric(participant))

ggplot(m_re_participant, aes(lv, .epred, colour = pthn, fill = pthn)) +
  facet_wrap(~participant) +
  geom_line(aes(group = interaction(pthn, participant)), size = 0.35) +
  geom_text(
    data = distinct(m_re_participant, participant), 
    aes(x = -1, y = 1, label = participant),
    size = 1.5, vjust = 1, hjust = 0, inherit.aes = FALSE, 
    position = position_nudge(x = 0.2),
    show.legend = FALSE
  ) +
  labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
       title = "Participant-level predictions") +
  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks = element_blank(),
    strip.text = element_blank(),
    strip.background = element_blank(),
    panel.spacing = unit(0.1, "lines")
  )

```

# Appendix 2: accuracy by item

## English-Spanish group

```{r appendix-accuracy-engspa, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.height = 7, fig.width = 4}

acc_engspa <- responses %>%
  filter(group == "ENG-SPA") %>% 
  group_by(word) %>%
  summarise(
    correct_sum = sum(correct, na.rm = TRUE),
    correct_n = sum(!is.na(correct), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  left_join(select(filter(stimuli, group=="ENG-SPA"), word = word1, word2, ipa1, ipa2)) %>% 
  drop_na(word2) %>% 
  rowwise() %>% 
  mutate(
    word = paste0(word, " /", ipa1, "/"),
    word2 = paste0(word2, " /", ipa2, "/"),
    correct_prop = prop_adj(correct_sum, correct_n),
    correct_se = prop_adj_se(correct_sum, correct_n),
    correct_ci_lower = correct_prop + qnorm((1-.95)/2)*correct_se,
    correct_ci_upper = correct_prop + qnorm(1-(1-.95)/2)*correct_se,
    correct_ci_lower = ifelse(correct_ci_lower < 0, 0, correct_ci_lower),
    correct_ci_upper = ifelse(correct_ci_upper > 1, 1, correct_ci_upper)
  ) %>% 
  ungroup() %>% 
  arrange(-correct_prop) %>% 
  select(word, word2, correct_prop, correct_se, correct_ci_lower, correct_ci_upper)

if ((nrow(acc_engspa) %% 2) > 0) {
  acc_engspa <- add_row(
    acc_engspa,
    data.frame(
      word = NA, word2 = NA, correct_prop = NA, correct_se = NA, 
      correct_ci_lower = NA, correct_ci_upper = NA
    )
  )
}

acc_engspa <- mutate(
  acc_engspa, 
  row = 1:nrow(acc_engspa), 
  col = ifelse(row <= nrow(acc_engspa)/2, 1, 2)
)

tab_engspa <- cbind(
  rename_all(acc_engspa[acc_engspa$col==1,], ~paste0(., "_1")), 
  rename_all(acc_engspa[acc_engspa$col==2,], ~paste0(., "_2"))
) %>% 
  select(-c(col_1, col_2, row_1, row_2)) %>% 
  as_tibble() %>% 
  gt() %>% 
  tab_header(
    title = "Table A.1",
    subtitle = "Proportion of correct translation by item in the English-Spanish group"
  ) %>% 
  tab_source_note("Words presented to participants in the English-Spanish group are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis. Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word. Accuracy is plotted separately for each group.") %>% 
  tab_spanner("(1)", ends_with("_1")) %>% 
  tab_spanner("(2)", ends_with("_2")) %>% 
  cols_merge(5:6, pattern = "[{1}, {2}]") %>% 
  cols_merge(11:12, pattern = "[{1}, {2}]") %>%

  fmt_percent(3:6, decimals = 0) %>% 
  fmt_percent(9:12, decimals = 0)  %>%

  cols_align("left") %>% 
  cols_label(
    word_1 = "Word",
    word2_1 = "Translation",
    correct_prop_1 = "Accuracy",
    correct_se_1 = md("*SE*"),
    correct_ci_lower_1 = md("*CI*<sub>95%</sub>"),
    word_2 = "Word",
    word2_2 = "Translation",
    correct_prop_2 = "Accuracy",
    correct_se_2 = md("*SE*"),
    correct_ci_lower_2 = md("*CI*<sub>95%</sub>")
  ) %>% 
  tab_style(
    style = cell_borders(sides = c("left", "right"), color = "white"),
    locations = list(
      cells_body(), cells_row_groups(),
      cells_stub(), 
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "small"), cell_borders(sides = "all", color = "white")),
    locations = cells_stub()
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "small", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = c(1:12))
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:6)
  ) %>% 
  tab_style(
    cell_text(weight = "bold"), cells_body(columns = c(1, 7))
  ) %>% 
  tab_style(
    cell_text(size = "small"), 
    cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_style(cell_text(size = "small"), cells_body(columns = everything(), rows = everything())) %>%   tab_options(data_row.padding = px(1)) %>% 
  gtsave(here("Figures", "accuracy_engspa.png"))

include_graphics(here("Figures", "accuracy_engspa.png"))
```

## English-Catalan group

```{r appendix-accuracy-engcat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.height = 7, fig.width = 4}

acc_engcat <- responses %>%
  filter(group == "ENG-CAT") %>% 
  group_by(word) %>%
  summarise(
    correct_sum = sum(correct, na.rm = TRUE),
    correct_n = sum(!is.na(correct), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  left_join(select(filter(stimuli, group=="ENG-CAT"), word = word1, word2, ipa1, ipa2)) %>% 
  drop_na(word2) %>% 
  rowwise() %>% 
  mutate(
    word = paste0(word, " /", ipa1, "/"),
    word2 = paste0(word2, " /", ipa2, "/"),
    correct_prop = prop_adj(correct_sum, correct_n),
    correct_se = prop_adj_se(correct_sum, correct_n),
    correct_ci_lower = correct_prop + qnorm((1-.95)/2)*correct_se,
    correct_ci_upper = correct_prop + qnorm(1-(1-.95)/2)*correct_se,
    correct_ci_lower = ifelse(correct_ci_lower < 0, 0, correct_ci_lower),
    correct_ci_upper = ifelse(correct_ci_upper > 1, 1, correct_ci_upper)
  ) %>% 
  ungroup() %>% 
  arrange(-correct_prop) %>% 
  select(word, word2, correct_prop, correct_se, correct_ci_lower, correct_ci_upper)

if ((nrow(acc_engcat) %% 2) > 0) {
  acc_engcat <- add_row(
    acc_engcat,
    data.frame(
      word = NA, word2 = NA, correct_prop = NA, correct_se = NA, 
      correct_ci_lower = NA, correct_ci_upper = NA
    )
  )
}

acc_engcat <- mutate(acc_engcat, row = 1:nrow(acc_engcat), col = ifelse(row <= nrow(acc_engcat)/2, 1, 2))

tab_engcat <- cbind(
  rename_all(acc_engcat[acc_engcat$col==1,], ~paste0(., "_1")), 
  rename_all(acc_engcat[acc_engcat$col==2,], ~paste0(., "_2"))
) %>% 
  select(-c(col_1, col_2, row_1, row_2)) %>% 
  as_tibble() %>% 
  gt() %>% 
  tab_header(
    title = "Table A.2",
    subtitle = "Proportion of correct translation by item in the English-Catalan group"
  ) %>% 
  tab_source_note("Words presented to participants in the English-Catalan group are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis. Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word. Accuracy is plotted separately for each group.") %>% 
  tab_spanner("(1)", ends_with("_1")) %>% 
  tab_spanner("(2)", ends_with("_2")) %>% 
  cols_merge(5:6, pattern = "[{1}, {2}]") %>% 
  cols_merge(11:12, pattern = "[{1}, {2}]") %>%

  fmt_percent(3:6, decimals = 0) %>% 
  fmt_percent(9:12, decimals = 0)  %>%

  cols_align("left") %>% 
  cols_label(
    word_1 = "Word",
    word2_1 = "Translation",
    correct_prop_1 = "Accuracy",
    correct_se_1 = md("*SE*"),
    correct_ci_lower_1 = md("*CI*<sub>95%</sub>"),
    word_2 = "Word",
    word2_2 = "Translation",
    correct_prop_2 = "Accuracy",
    correct_se_2 = md("*SE*"),
    correct_ci_lower_2 = md("*CI*<sub>95%</sub>")
  ) %>% 
  tab_style(
    style = cell_borders(sides = c("left", "right"), color = "white"),
    locations = list(
      cells_body(), cells_row_groups(),
      cells_stub(), 
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "small"), cell_borders(sides = "all", color = "white")),
    locations = cells_stub()
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "small", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = c(1:12))
  ) %>% 
  tab_style(
    cell_text(weight = "bold"), cells_body(columns = c(1, 7))
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:6)
  ) %>% 
  tab_style(
    cell_text(size = "small"), 
    cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_style(cell_text(size = "small"), cells_body(columns = everything(), rows = everything())) %>%   tab_options(data_row.padding = px(1)) %>% 
  gtsave(here("Figures", "accuracy_engcat.png"))

include_graphics(here("Figures", "accuracy_engcat.png"))

```

## Spanish-Catalan group

```{r appendix-accuracy-spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.height = 7, fig.width = 4}

acc_spacat <- responses %>%
  filter(group == "SPA-CAT") %>% 
  group_by(word) %>%
  summarise(
    correct_sum = sum(correct, na.rm = TRUE),
    correct_n = sum(!is.na(correct), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  left_join(select(filter(stimuli, group=="SPA-CAT"), word = word1, word2, ipa1, ipa2)) %>% 
  drop_na(word2) %>% 
  rowwise() %>% 
  mutate(
    word = paste0(word, " /", ipa1, "/"),
    word2 = paste0(word2, " /", ipa2, "/"),
    correct_prop = prop_adj(correct_sum, correct_n),
    correct_se = prop_adj_se(correct_sum, correct_n),
    correct_ci_lower = correct_prop + qnorm((1-.95)/2)*correct_se,
    correct_ci_upper = correct_prop + qnorm(1-(1-.95)/2)*correct_se,
    correct_ci_lower = ifelse(correct_ci_lower < 0, 0, correct_ci_lower),
    correct_ci_upper = ifelse(correct_ci_upper > 1, 1, correct_ci_upper)
  ) %>% 
  ungroup() %>% 
  arrange(-correct_prop) %>% 
  select(word, word2, correct_prop, correct_se, correct_ci_lower, correct_ci_upper)

if ((nrow(acc_spacat) %% 2) > 0) {
  acc_spacat <- add_row(
    acc_spacat,
    data.frame(
      word = NA, word2 = NA, correct_prop = NA, correct_se = NA, 
      correct_ci_lower = NA, correct_ci_upper = NA
    )
  )
}

acc_spacat <- mutate(acc_spacat, row = 1:nrow(acc_spacat), col = ifelse(row <= nrow(acc_spacat)/2, 1, 2))

tab_spacat <- cbind(
  rename_all(acc_spacat[acc_spacat$col==1,], ~paste0(., "_1")), 
  rename_all(acc_spacat[acc_spacat$col==2,], ~paste0(., "_2"))
) %>% 
  select(-c(col_1, col_2, row_1, row_2)) %>% 
  as_tibble() %>% 
  gt() %>% 
  tab_header(
    title = "Table A.3",
    subtitle = "Proportion of correct translation by item in the Spanish-Catalan group"
  ) %>% 
  tab_source_note("Words presented to participants in the Spanish-Catalan group are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis. Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word. Accuracy is plotted separately for each group.") %>% 
  tab_spanner("(1)", ends_with("_1")) %>% 
  tab_spanner("(2)", ends_with("_2")) %>% 
  cols_merge(5:6, pattern = "[{1}, {2}]") %>% 
  cols_merge(11:12, pattern = "[{1}, {2}]") %>%

  fmt_percent(3:6, decimals = 0) %>% 
  fmt_percent(9:12, decimals = 0)  %>%

  cols_align("left") %>% 
  cols_label(
    word_1 = "Word",
    word2_1 = "Translation",
    correct_prop_1 = "Accuracy",
    correct_se_1 = md("*SE*"),
    correct_ci_lower_1 = md("*CI*<sub>95%</sub>"),
    word_2 = "Word",
    word2_2 = "Translation",
    correct_prop_2 = "Accuracy",
    correct_se_2 = md("*SE*"),
    correct_ci_lower_2 = md("*CI*<sub>95%</sub>")
  ) %>% 
  tab_style(
    style = cell_borders(sides = c("left", "right"), color = "white"),
    locations = list(
      cells_body(), cells_row_groups(),
      cells_stub(), 
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "small"), cell_borders(sides = "all", color = "white")),
    locations = cells_stub()
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "small", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = c(1:12))
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:6)
  ) %>% 
  tab_style(
    cell_text(weight = "bold"), cells_body(columns = c(1, 7))
  ) %>% 
  tab_style(
    cell_text(size = "small"), 
    cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_options(data_row.padding = px(1)) %>% 
  gtsave(here("Figures", "accuracy_spacat.png"))

include_graphics(here("Figures", "accuracy_spacat.png"))

```