### 2021-12-20 docs/00_stimuli.Rmd
Reorganise repo

---
title: "Stimuli"
output:
  md_document:
    variant: markdown_github
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "80%",
  results = "asis",
  dpi = 300,
  dev.args = list(png = list(type = "cairo"))
)

options(
  knitr.kable.NA = '-',
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.continuous.fill = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  ),
  ggplot2.continuous.colour = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  )
)
```

```{r prepare}
# import data
tar_load_globals()
tar_load(stimuli)

theme_set(theme_custom())
```

# Overview

Participants in the ENG-SPA listened to `r nrow(stimuli[stimuli$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials. Five trials in each condition were practice trials and were excluded from analyses.

```{r stimuli_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
stimuli_coded <- mutate(stimuli, test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"))
```

# Lexical frequency

```{r stimuli_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  distinct(test_language, word1, .keep_all = TRUE) %>% 
  left_join(count(., test_language, name = "n_total")) %>% 
  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
  ggplot(aes(frequency_zipf, fill = test_language)) +
  facet_wrap(~test_language, ncol = 1) +
  geom_histogram(bins = 25, colour = "white") +
  labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language")  +
  scale_y_continuous(breaks = seq(0, 12, 2)) +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

```

# PTHN

```{r stimuli_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  left_join(count(., test_language, name = "n_total")) %>% 
  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
  count(test_language, pthn) %>% 
  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
  facet_wrap(~test_language, ncol = 1) +
  geom_col(position = position_dodge()) +
  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

```

# Phonological similarity

```{r stimuli_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  left_join(count(., group, name = "n_total")) %>% 
  mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
  ggplot(aes(group, lv, fill = group, colour = group)) +
  geom_violin() +
  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
  labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

```