### 2021-12-20 docs/02_data-analysis.Rmd

---
title: "Data analysis"
output:
  md_document:
    variant: markdown_github
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "80%",
  results = "asis",
  dpi = 300,
  dev.args = list(png = list(type = "cairo"))
)
options(
  knitr.kable.NA = '-',
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.continuous.fill = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  ),
  ggplot2.continuous.colour = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  )
)
```

```{r prepare}
# import data
tar_load_globals()

theme_set(theme_custom())
```

# Model details

We fitted a Bayesian multilevel regression model using the R package `brms` [@burkner2017brms], with correct responses (`correct`, 0 is incorrect response, 1 if correct response) as the response variable with a Bernoulli distribution and a logit link function. We modelled the probability of an average participant providing a correct response, conditional to a set of predictors (aka. fixed effects) and their interactions:

* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database. Standardised before entering the regression model.
* Phonological similarity (`lv`, 0-1): Inverse Levenshtein distance between the IPA transcription of the presented word and its correct translation in the target language, calculated using the `stringdist` function of the `stringdist` package [@loo2014stringdist]. Standardised before entering the regression model.
* Shared phonological onset (`onset`, Same/Different): Whether the presented and the target word share phonological onset, as judged by experimenters, and sum coded as $Different = -0.5$ and $Same = +0.5$.

We added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model. We specified random intercepts slopes for all fixed effects within the `participant` grouping variable.

We performed multiple imputation via predictive mean matching using the `mice` function of the `mice` package [@van2011mice] to impute missing values in the response variable or the predictors.

The model was implemented using the following formula (`fit_3`):

$$
\begin{align}

&\textbf{Likelihood}  \\
y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\

&\textbf{Parameters}  \\

logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
& \beta_{1[p]} ~ Frequency_{i} ~ + \\
& \beta_{2[p]} ~ PTHN_i ~ + \\
& \beta_{3[p]} ~ Similarity_i ~ + \\
& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\

\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
&& \text{[participant-level coefficients]} \\ \\

&\textbf{Prior}  \\

\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\

\end{align}
$$

The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 iterations each (1,000 warm-ups iterations). The output of this model is the approximated posterior distribution of all the parameters estimated, therefore indicating the probability of each combination of values of the parameters, conditional to the observed data.

We compared this (extended model) to other models dropping one predictor at a time using Leave-one-out cross-validation (LOO-CV), using the `loo` and `loo_compare` functions of the `brms` package. We compared our extended model against the following models:

* fit_0: `~ 1 + (1 | participant) + (1 | word)`
* fit_1: `~ 1 + frequency_zipf + (1 + frequency_zipf | participant) + (1 | word)`
* fit_2: `~ 1 + frequency_zipf + pthn + (1 + frequency_zipf + pthn | participant) + (1 | word)`
* fit_3: `~ 1 + frequency_zipf + pthn + lv + (1 + frequency_zipf + pthn + lv | participant) + (1 | word)`
* fit_4: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_5: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_6: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + group:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`

# Model implementation

## Stan code

```stan stan_model
// generated with brms 2.16.1
functions {
 /* compute correlated group-level effects
  * Args: 
  *   z: matrix of unscaled group-level effects
  *   SD: vector of standard deviation parameters
  *   L: cholesky factor correlation matrix
  * Returns: 
  *   matrix of scaled group-level effects
  */ 
  matrix scale_r_cor(matrix z, vector SD, matrix L) {
    // r is stored in another dimension order than z
    return transpose(diag_pre_multiply(SD, L) * z);
  }
}
data {
  int<lower=1> N;  // total number of observations
  int Y[N];  // response variable
  int<lower=1> K;  // number of population-level effects
  matrix[N, K] X;  // population-level design matrix
  // data for group-level effects of ID 1
  int<lower=1> N_1;  // number of grouping levels
  int<lower=1> M_1;  // number of coefficients per level
  int<lower=1> J_1[N];  // grouping indicator per observation
  // group-level predictor values
  vector[N] Z_1_1;
  vector[N] Z_1_2;
  vector[N] Z_1_3;
  vector[N] Z_1_4;
  vector[N] Z_1_5;
  vector[N] Z_1_6;
  vector[N] Z_1_7;
  int<lower=1> NC_1;  // number of group-level correlations
  // data for group-level effects of ID 2
  int<lower=1> N_2;  // number of grouping levels
  int<lower=1> M_2;  // number of coefficients per level
  int<lower=1> J_2[N];  // grouping indicator per observation
  // group-level predictor values
  vector[N] Z_2_1;
  int prior_only;  // should the likelihood be ignored?
}
transformed data {
  int Kc = K - 1;
  matrix[N, Kc] Xc;  // centered version of X without an intercept
  vector[Kc] means_X;  // column means of X before centering
  for (i in 2:K) {
    means_X[i - 1] = mean(X[, i]);
    Xc[, i - 1] = X[, i] - means_X[i - 1];
  }
}
parameters {
  vector[Kc] b;  // population-level effects
  real Intercept;  // temporary intercept for centered predictors
  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
  matrix[M_1, N_1] z_1;  // standardized group-level effects
  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
  vector[N_2] z_2[M_2];  // standardized group-level effects
}
transformed parameters {
  matrix[N_1, M_1] r_1;  // actual group-level effects
  // using vectors speeds up indexing in loops
  vector[N_1] r_1_1;
  vector[N_1] r_1_2;
  vector[N_1] r_1_3;
  vector[N_1] r_1_4;
  vector[N_1] r_1_5;
  vector[N_1] r_1_6;
  vector[N_1] r_1_7;
  vector[N_2] r_2_1;  // actual group-level effects
  // compute actual group-level effects
  r_1 = scale_r_cor(z_1, sd_1, L_1);
  r_1_1 = r_1[, 1];
  r_1_2 = r_1[, 2];
  r_1_3 = r_1[, 3];
  r_1_4 = r_1[, 4];
  r_1_5 = r_1[, 5];
  r_1_6 = r_1[, 6];
  r_1_7 = r_1[, 7];
  r_2_1 = (sd_2[1] * (z_2[1]));
}
model {
  // likelihood including constants
  if (!prior_only) {
    // initialize linear predictor term
    vector[N] mu = Intercept + rep_vector(0.0, N);
    for (n in 1:N) {
      // add more terms to the linear predictor
      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n];
    }
    target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
  }
  // priors including constants
  target += normal_lpdf(b | 0, 0.1);
  target += normal_lpdf(Intercept | 0, 0.1);
  target += cauchy_lpdf(sd_1 | 0, 0.1)
    - 7 * cauchy_lccdf(0 | 0, 0.1);
  target += std_normal_lpdf(to_vector(z_1));
  target += lkj_corr_cholesky_lpdf(L_1 | 8);
  target += cauchy_lpdf(sd_2 | 0, 0.1)
    - 1 * cauchy_lccdf(0 | 0, 0.1);
  target += std_normal_lpdf(z_2[1]);
}
generated quantities {
  // actual population-level intercept
  real b_Intercept = Intercept - dot_product(means_X, b);
  // compute group-level correlations
  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
  vector<lower=-1,upper=1>[NC_1] cor_1;
  // additionally sample draws from priors
  real prior_b = normal_rng(0,0.1);
  real prior_Intercept = normal_rng(0,0.1);
  real prior_sd_1 = cauchy_rng(0,0.1);
  real prior_cor_1 = lkj_corr_rng(M_1,8)[1, 2];
  real prior_sd_2 = cauchy_rng(0,0.1);
  // extract upper diagonal of correlation matrix
  for (k in 1:M_1) {
    for (j in 1:(k - 1)) {
      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
    }
  }
  // use rejection sampling for truncated priors
  while (prior_sd_1 < 0) {
    prior_sd_1 = cauchy_rng(0,0.1);
  }
  while (prior_sd_2 < 0) {
    prior_sd_2 = cauchy_rng(0,0.1);
  }
}
```