### 2022-02-18 README.Rmd
Render README as target

---
output: github_document
clean: true
---

# TranslationElicitation

More info in the Open Science Framework repository.

## How to make this repository work on your local machine

### 1. Downloading the repository

You can either download the project as a .zip file by clocking on the green button that says "Code", or if you feel comfortable with Git, you can clone the repository from your console running:

```bash
git clone https://github.com/gongcastro/translation-elicitation.git
```

### 2. Setting up R packages with renv

We used the R package renv to keep track of the dependencies of the project. renv allows to install and update the R packages necessary to run the project in a project-specific environment. Packages installed using renv in this project will only be installed or updated in the context of this project. The R packages you had installed for other projects will remain unaffected.

1) Open this project in a new RStudio session (open the project folder and click on the `translation-elicitation.Rproj` file). 
2) In the R console, run the following lines of code:

```r
# install.packages("renv") # in case yopu need to install renv
renv::activate() # this will create the renv infrastructure if missing
renv::restore() # this will install/update/move the necessary R pakcages into this project
```

### 3. Running the code with targets

Restart your R session. The code in the .Rprofile file should run at start, and define the function `make()`. You should be able to run `make()` in your console. This will run all lines of code in the right order, and will generate all objects and outputs (e.g., data frames) used in the project, and render the manuscript.

Here are some details about the code, in case you want to dive into it. We use the targets package to keep track of what code should be run, and in what order. The `_targets.R` script contains the main commands in the workflow. Each command is indicated inside of a call to the `tar_target()` function, and is called a *target*. For readability, we have defined most of the functions used inside the targets on separate scripts, hosted in the R/ folder of the project. Take a look at them if you want to see what exactly what happens inside those function.

You can import the objects generated by the targets using the `tar_load()` function. First, load all necessary R packages running `tar_load_globals()`. Then, if you have previously run `make()`, you can run `tar_load(responses)` to get the main data set or `tar_load(model_fits)` to inspect the models, for instance. You don't have to assign the call to the `tar_load()` function to a variable: the object will be directly save as a variable in your workspace.

```{r targets, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

targets::tar_visnetwork()

```

## How is this repository organised?

This repository is organised as follows:

* **data**: raw, processed, and coded data from the behavioural task
* **img**: figures resulting from the scripts
* **docs**: lab notes
* **manuscript**: Rmarkdown files and outputs of the manuscript (generated using the {papaja} R package)
* **results**: outputs of the models.
* **scripts**: R code used to process and analyse the data, and Stan code used to fit the Bayesian models (called via the R scripts).
* **stimuli**: Stimuli and trial lists used in the behavioural task.

The file `.gitignore` indicates Git what files/subfolders must *not* be kept track of (for privacy and storage limit, e.g., audios).

The experimental set up and code (PsychoPy/Pavlovia) are stored in the following GitLab repositories:

* Version in Catalan for **Spanish speakers**: https://gitlab.pavlovia.org/gongcastro/translationelicitationspa
* Version in Spanish or Catalan for **English speakers**: https://gitlab.pavlovia.org/SiowSerene/translationelicitation_eng