### 2024-04-02 docs/index.qmd
Make report in Quarto

---
title: "Translation Elicitation"
subtitle: "Lab notes"
format: html
---

```{r}
#| label: setup
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggdist)

tar_load(stimuli)
```

## Stimuli

### Lexical frequency

```{r}
stimuli |>
    ggplot(aes(group, freq_2, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    labs(x = "Average cross-linguistic phonological similarity",
         y = "Lexical frequency (log10)") +
    scale_y_continuous(transform = "log10")
```

### Cross-linguistic phonological similarity (cognateness)

```{r}
stimuli |>
    ggplot(aes(lv, group = NA)) +
    facet_wrap(~group) +
    geom_histogram() +
    labs(x = "Average cross-linguistic phonological similarity",
         y = "Lexical frequency (log10)")
```

### Cross-linguistic neighbourhood density

```{r}
#| label: fig-clnd
#| fig-cap: "Cross-language neighbourhood density"
stimuli |>
    mutate(total = n(), .by = group) |> 
    count(group, neigh_n, total) |> 
    mutate(n = n/total) |> 
    ggplot(aes(neigh_n, n)) +
    facet_wrap(~group) +
    geom_col(colour = "white") +
    labs(x = "Number of cross-language phonological neighbours",
         y = "Proportion of items")
```

### Cross-linguistic neighbourhood similarity

```{r}
#| label: fig-clnd
#| fig-cap: "Cross-language neighbourhood similarity"
stimuli |>
    ggplot(aes(group, neigh_sim, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    labs(x = "Group",
         y = "Average cross-linguistic phonological similarity")
```

### Audio duration

```{r}
#| label: fig-clnd
#| fig-cap: "Cross-language neighbourhood similarity"
stimuli |>
    ggplot(aes(group, duration, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    labs(x = "Audio duration (s)",
         y = "Average cross-linguistic phonological similarity")
```