### 2024-05-02 docs/index.qmd
Update report

---
title: "Translation Elicitation"
subtitle: "Lab notes"
format: html
---

```{r}
#| label: setup
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gt)
library(gtExtras)
library(patchwork)
library(gghighlight)
library(tidybayes)

targets::tar_config_set(
    store = here::here("_targets"),
    script = here::here("_targets.R")
)

tar_load(stimuli)
tar_load(exp_participants)
tar_load(exp_responses)
tar_load(quest_participants)
tar_load(quest_responses)
tar_load(dataset_1)
tar_load(dataset_2)
tar_load(dataset_3)
tar_load(fit_1)

theme_set(theme_ggdist() + 
              theme(panel.grid.major.y = element_line(
                  linetype = "dotted",
                  colour = "grey")))

# proportion adjusted from boundary values (Gelman, Hill & Vehtari, 2020)
prop_adj <- function(x, n){
    e <- (x+2)/(n+4)
    return(e)
}

# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
prop_adj_se <- function(x, n) {
    e <- (x+2)/(n+4)
    se <- sqrt(e*(1-e)/(n+4))
    return(se)
}

# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
prop_adj_ci <- function(x, n, .width = 0.95) {
    e <- (x+2)/(n+4)
    se <- sqrt(e*(1-e)/(n+4))
    ci <-  e + qnorm(c((1-.width)/2, (1-(1-.width)/2)))*se
    ci[1] <- ifelse(ci[1]<0, 0, ci[1]) # truncate at 0
    ci[2] <- ifelse(ci[2]>1, 1, ci[2]) # truncate at 1
    return(ci)
}
```

# Stimuli

## Audio duration

```{r}
#| label: fig-duration
#| fig-cap: "Cross-language neighbourhood similarity"
stimuli |>
    ggplot(aes(group, duration, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    stat_summary(fun.data = mean_se,
                 geom = "errorbar",
                 colour = "red",
                 linewidth = 1,
                 width = 0.1) +
    stat_summary(fun = mean,
                 geom = "point",
                 colour = "red",
                 size = 2) +
    labs(x = "Audio duration (s)",
         y = "Average cross-linguistic phonological similarity")
```

### Lexical frequency

Lexical frequency in the TARGET language (English or Spanish).

```{r}
#| label: frequency-2
missing <- stimuli[is.na(stimuli$freq_zipf_2), "ipa_flat_1"]
```

Missing observations: `r paste0(missing, collapse = ", ")`.

```{r}
stimuli |>
    drop_na(freq_zipf_2) |> 
    ggplot(aes(group, freq_zipf_2, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both",
              alpha = 0.5) +
    stat_summary(fun.data = mean_se, color = "red",
                 geom = "errorbar",
                 width = 0.5,
                 linewidth = 1) +
    labs(x = "Average cross-linguistic phonological similarity",
         y = "Lexical frequency (Zipf)")
```

## Phonological similarity between translation pairs (cognateness)

```{r}
stimuli |> 
    mutate(t_label = paste0(ipa_flat_1, "-", ipa_flat_2)) |> 
    ggplot(aes(group, lv)) +
    facet_wrap(~group, scales = "free_x") +
    stat_interval(size = 8) +
    # geom_text(data = stimuli |>
    #               mutate(t_label = paste0(ipa_flat_1, "-", ipa_flat_2)) |>
    #               top_n(10, lv),
    #           aes(label = t_label),
    #           position = position_nudge(x = 0.1),
    #           hjust = 0)
    # geom_point() +
    ggrepel::geom_text_repel(aes(label = t_label),
                             seed = 1234, size = 3,
                             direction = "y",
                             box.padding = unit(0.2, "lines"),
                             max.overlaps = 13,
                             alpha = 0.55,
                             segment.size = 0,
                             segment.curvature = 1/4) +
    labs(x = "Group",
         y = "Cognateness\n(Levenshtein similarity)",
         colour = "Proportion of translations") +
    scale_y_continuous(labels = scales::percent) +
    scale_color_brewer(palette = "Reds") +
    theme(legend.position = "top",
          axis.text.x = element_blank())
```

## Cross-linguistic neighbourhood density

```{r}
#| label: tbl-clnd
#| fig-cap: "Cross-language neighbourhood density"
stimuli |> 
    arrange(group, desc(neigh_n)) |> 
    group_by(group) |> 
    top_n(n = 10, neigh_n) |> 
    select(group, translation, avg_sim, neigh_n, neigh_lst) |> 
    gt(rowname_col = "translation",
       groupname_col = "group") |> 
    fmt_percent(avg_sim) |> 
    tab_style(cell_text(align = "left"), cells_body()) |> 
    cols_label(neigh_n = "N neighbours",
               avg_sim = "Average cross-language similarity",
               neigh_lst = "List of neighbours")
```

## Cross-linguistic neighbourhood density (higher frequency than translation)

```{r}
#| label: tbl-clnd
#| fig-cap: "Cross-language neighbourhood density (higher frequency than translation)"
stimuli |> 
    arrange(group, desc(neigh_n_h)) |> 
    group_by(group) |> 
    top_n(n = 10, neigh_n_h) |> 
    select(group, translation, avg_sim_h, neigh_n_h, neigh_lst_h) |> 
    gt(rowname_col = "translation",
       groupname_col = "group") |> 
    fmt_percent(avg_sim_h) |> 
    tab_style(cell_text(align = "left"), cells_body()) |> 
    cols_label(neigh_n_h = "N neighbours",
               avg_sim_h = "Average cross-language similarity",
               neigh_lst_h = "List of neighbours")
```

## Cross-linguistic neighbourhood similarity

```{r}
#| label: fig-clns
#| fig-cap: "Cross-language neighbourhood similarity"
stimuli |>
    ggplot(aes(group, avg_sim, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    stat_summary(fun.data = mean_se,
                 geom = "errorbar",
                 colour = "red",
                 linewidth = 3/4,
                 width = 0.1) +
    stat_summary(fun = mean,
                 geom = "point",
                 colour = "red",
                 size = 2) +
    labs(x = "Group",
         y = "Average cross-linguistic phonological similarity")
```

## Cognateness by average cross-linguistic similarity

```{r}
#| label: fig-cognateness-avg-sim
stimuli |> 
    ggplot(aes(lv, avg_sim)) +
    geom_rug(colour = "grey") +
    geom_point(alpha = 0.5,
               shape = 1,
               size = 2,
               stroke = 1) +
    geom_smooth(method = "lm",
                colour = "red",
                fill = "red") + 
    labs(x = "Cognateness (Levenshtein distance)",
         y = "Average cross-language similarity")
```

# Participants

```{r}
#| label: participants
tbl_data <- list(Experiment = exp_participants,
                 Questionnaire = quest_participants) |> 
    bind_rows(.id = "source") |> 
    summarise(n = n(),
              n_excluded = sum(!valid_participant),
              across(age, lst(mean, sd, min, max)),
              knows_l2 = sum(l_2 != "None"),
              l2_lst = lst(unique(l_2)[unique(l_2)!="None"]),
              .by = c(group, source))

tbl_data |> 
    gt(groupname_col = "source", rowname_col = "group") |> 
    fmt_number(matches("age")) |> 
    fmt_integer(matches("min|max")) |> 
    cols_merge_range(age_min, age_max) |> 
    tab_spanner("Age", matches("age")) |> 
    tab_spanner("Sample size", starts_with("n")) |> 
    tab_spanner("L2", matches("l2")) |> 
    cols_label(group = "Group",
               source = "Source",
               n = "N",
               n_excluded = "Excluded",
               age_mean = "M",
               age_sd = "SD",
               age_min = "Range",
               knows_l2 = "N knows L2?",
               l2_lst = "Which L2?") |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels()) |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(align = "left"),
              list(cells_body(), cells_column_labels()))
```

## Target language proficiency

```{r}
tbl_data <- list(Experiment = exp_participants,
                 Questionnaire = quest_participants) |> 
    bind_rows(.id = "source") |> 
    summarise(across(matches("writ|oral"),
                     lst(mean, sd, min, max)),
              .by = c(group, source)) |> 
    select(-matches("l_2"))

tbl_data |> 
    gt(rowname_col = "group", groupname_col = "source") |> 
    fmt_number(is.numeric) |> 
    fmt_integer(matches("min|max")) |> 
    fmt_missing(rows = everything(), columns = everything(),
                missing_text = "--") |> 
    cols_merge_range(cat_oral_comp_min, cat_oral_comp_max) |> 
    cols_merge_range(spa_oral_comp_min, spa_oral_comp_max) |> 
    cols_merge_range(cat_writ_prod_min, cat_writ_prod_max) |> 
    cols_merge_range(spa_writ_prod_min, spa_writ_prod_max) |> 
    tab_spanner("Catalan", matches("cat_")) |> 
    tab_spanner("Spanish", matches(("spa_"))) |> 
    tab_spanner("Oral comprehension", matches("oral")) |> 
    tab_spanner("Written production", matches("writ")) |> 
    cols_label(cat_oral_comp_mean = "M",
               spa_oral_comp_mean = "M",
               cat_writ_prod_mean = "M",
               spa_writ_prod_mean = "M",
               cat_oral_comp_sd = "SD",
               spa_oral_comp_sd = "SD",
               cat_writ_prod_sd = "SD",
               spa_writ_prod_sd = "SD",
               cat_oral_comp_min = "Range",
               spa_oral_comp_min = "Range",
               cat_writ_prod_min = "Range",
               spa_writ_prod_min = "Range") |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels()) |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(align = "left"),
              list(cells_body(), cells_column_labels()))
```

# Results

```{r}
#| label: tbl-dataset
sem <- function(x) mean(x) / (sqrt(length(x)))

dataset_2 |> 
    add_count(source, group, participant_id, name = "trials") |> 
    summarise(n_trials = n(),
              correct = sum(correct),
              .by = c(group, source, participant_id)) |> 
    mutate(correct = prop_adj(correct, n_trials)) |> 
    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
              across(n_trials, lst(mean, sum, sd, min, max)),
              n_participants = n_distinct(participant_id),
              .by = c(group, source)) |> 
    relocate(n_participants, matches("n_trials")) |> 
    gt(rowname_col = "group",
       groupname_col = "source") |> 
    fmt_number(is.numeric, decimals = 2) |> 
    gtExtras::gt_plt_dist(correct_list, type = "density") |> 
    fmt_integer(c(matches("sum|min|max"), n_participants)) |> 
    fmt_number(matches("correct"), scale_by = 100) |> 
    tab_spanner("Valid trials", matches("n_trials")) |> 
    tab_spanner("Accuracy (%)", matches("correct")) |> 
    cols_merge_range(n_trials_min, n_trials_max) |> 
    cols_merge_range(correct_min, correct_max) |> 
    cols_label(n_trials_sum = "N trials",
               n_participants = "N",
               n_trials_mean = "Mean",
               n_trials_sd = "SD",
               n_trials_min = "Range",
               correct_mean = "Mean",
               correct_sd = "SD",
               correct_sem = "SE",
               correct_min = "Range",
               correct_list = "") |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels()) |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(align = "left"),
              list(cells_body(), cells_column_labels()))
```

## Questionnaire results

```{r}
quest_responses |> 
    summarise(n_trials = n(),
              correct = sum(correct),
              .by = c(group, confidence, participant_id)) |> 
    mutate(correct = prop_adj(correct, n_trials)) |> 
    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
              .by = c(group, confidence)) |> 
    ggplot(aes(confidence, correct_mean,
               colour = group,
               ymin = correct_mean - correct_sem,
               ymax = correct_mean + correct_sem)) +
    geom_errorbar(width = 0.5,
                  linewidth = 3/4,
                  position = position_dodge(width = 0.5)) +
    geom_point(size = 2,
               position = position_dodge(width = 0.5)) +
    scale_y_continuous(limits = c(0, 1)) +
    labs(x = "Reported confidence (1-5)",
         y = "Accuracy",
         colour = "Group") +

    quest_responses |> 
    summarise(n_trials = n(),
              correct = sum(correct),
              .by = c(group, knowledge, participant_id)) |> 
    mutate(correct = prop_adj(correct, n_trials)) |> 
    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
              .by = c(group, knowledge)) |> 
    mutate(knowledge = if_else(knowledge, "Yes", "No")) |> 
    ggplot(aes(knowledge, correct_mean,
               colour = group,
               ymin = correct_mean - correct_sem,
               ymax = correct_mean + correct_sem)) +
    geom_errorbar(width = 0.25,
                  linewidth = 3/4,
                  position = position_dodge(width = 0.5)) +
    geom_point(size = 2,
               position = position_dodge(width = 0.5)) +
    scale_y_continuous(limits = c(0, 1)) +
    labs(x = "Reported word knowledge",
         y = "Accuracy",
         colour = "Group") +
    theme(axis.text.y = element_blank(),
          axis.title.y = element_blank()) +

    quest_responses |> 
    mutate(knowledge = if_else(knowledge, "Yes", "No")) |> 
    summarise(across(confidence, lst(mean, sem)),
              .by = c(group, knowledge)) |> 
    ggplot(aes(knowledge, confidence_mean,
               colour = group,
               ymin = confidence_mean - confidence_sem,
               ymax = confidence_mean + confidence_sem)) +
    geom_errorbar(width = 0.25,
                  linewidth = 3/4,
                  position = position_dodge(width = 0.5)) +
    geom_point(size = 2,
               position = position_dodge(width = 0.5)) +
    labs(x = "Reported word knowledge",
         y = "Resported confidence (1-5)",
         colour = "Group") +
    scale_y_continuous(labels = scales::percent) +
    plot_layout(nrow = 1,
                guides = "collect") &
    theme(legend.position = "top")
```

```{r}
#| label: fig-results-lv
#| fig-cap: "Accuracy by cognateness."
dataset_2 |> 
    mutate(lv_std = cut_interval(lv_std, 10, labels = FALSE),
           group = factor(group, levels = c("cat-ENG",
                                            "spa-ENG",
                                            "cat-SPA"))) |>
    summarise(n_trials = n(),
              correct = sum(correct),
              .by = c(group, lv_std, participant_id)) |> 
    mutate(correct = prop_adj(correct, n_trials)) |> 
    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
              .by = c(group, lv_std)) |> 
    ggplot(aes(lv_std, correct_mean, colour = group,
               ymin = correct_mean-correct_sem,
               ymax = correct_mean+correct_sem)) +
    facet_wrap(~group) +
    geom_errorbar(width = 0.25,
                  linewidth = 3/4,
                  position = position_dodge(width = 0.25)) +
    geom_point(size = 2,
                  position = position_dodge(width = 0.25)) +
    labs(x = "SD Cognateness (Levenshtein distance)",
         y = "Accuracy",
         colour = "Group",
         fill = "Group") +
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "none")
```

```{r}
#| label: fig-results-avg_sim
#| fig-cap: "Accuracy by average cross-language similarity"
dataset_2 |> 
    mutate(avg_sim_std = cut_number(avg_sim_std, 10, labels = FALSE)) |>
    summarise(n_trials = n(),
              correct = sum(correct),
              .by = c(group, avg_sim_std, participant_id)) |> 
    mutate(correct = prop_adj(correct, n_trials)) |> 
    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
              .by = c(group, avg_sim_std)) |> 
    ggplot(aes(avg_sim_std, correct_mean, colour = group,
               ymin = correct_mean-correct_sem,
               ymax = correct_mean+correct_sem)) +
    facet_wrap(~group) +
    geom_errorbar(width = 0.25,
                  linewidth = 3/4) +
    geom_point(size = 2) +
    labs(x = "SD Average cross-language similarity",
         y = "Accuracy",
         colour = "Group",
         fill = "Group") +
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "none")
```

```{r}
#| label: fig-results-avg_sim
#| fig-cap: "Accuracy by average cross-language similarity"
dataset_2 |> 
    summarise(n_trials = n(),
              correct = sum(correct),
              .by = c(group, neigh_n, participant_id)) |> 
    mutate(correct = prop_adj(correct, n_trials)) |> 
    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
              .by = c(group, neigh_n)) |> 
    ggplot(aes(neigh_n, correct_mean, colour = group,
               ymin = correct_mean-correct_sem,
               ymax = correct_mean+correct_sem)) +
    geom_point(size = 2) +
    geom_smooth(aes(group = group),
                method = "glm", 
                method.args = list(family = "binomial"), 
                se = FALSE) +
    labs(x = "Cognateness (Levenshtein distance)",
         y = "Accuracy",
         colour = "Group",
         fill = "Group")
```
## Model outputs

### Analysis 1

```{r}
nd <- expand_grid(lv_std = seq(-3, 3, 0.1),
                  avg_sim_h_std = 0,
                  freq_zipf_2_std = 0)

epreds <- add_epred_draws(nd, fit_0, re_formula = NA) |> 
    # mean_qi() |> 
    mutate(avg_sim_h_std = as.factor(avg_sim_h_std))

dataset_1 |> 
    summarise(n_trials = n(),
              correct = sum(correct),
              .by = c(lv_std, participant_id)) |> 
    mutate(correct = prop_adj(correct, n_trials)) |> 
    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
              .by = c(lv_std)) |> 
    ggplot(aes(lv_std, .epred)) +
    stat_lineribbon(data = epreds) +
    geom_point(aes(y = correct_mean)) +
    scale_fill_brewer(palette = "Reds")
```

# Appendix

## Appendix 1: List of trials

### Catalan-English (cat-ENG)

```{r}
#| label: tbl-app-stim-cateng
#| tbl-cap: "Catalan-English stimuli"
tar_load(trial_list)

trial_list_lst <- trial_list |> 
    split(trial_list$group) |>
    janitor::clean_names() |> 
    map(select, matches("word_|ipa|sampa"))

trial_list_lst$cat_eng |> 
    gt() |> 
    tab_spanner("Presented word (1)", ends_with("_1")) |> 
    tab_spanner("Presented word (2)", ends_with("_2")) |> 
    cols_label(word_1 = "Orthography",
               ipa_flat_1 = "IPA",
               sampa_1 = "X-SAMPA",
               word_2 = "Orthography",
               ipa_flat_2 = "IPA",
               sampa_2 = "X-SAMPA")
```

### Spanish-English (spa-ENG)

```{r}
#| label: tbl-app-stim-cateng
#| tbl-cap: "Spanish-English stimuli"
trial_list_lst$spa_eng |> 
    gt() |> 
    tab_spanner("Presented word (1)", ends_with("_1")) |> 
    tab_spanner("Presented word (2)", ends_with("_2")) |> 
    cols_label(word_1 = "Orthography",
               ipa_flat_1 = "IPA",
               sampa_1 = "X-SAMPA",
               word_2 = "Orthography",
               ipa_flat_2 = "IPA",
               sampa_2 = "X-SAMPA")
```

### Catalan-Spanish (cat-SPA)

```{r}
#| label: tbl-app-stim-catspa
#| tbl-cap: "Catalan-Spanish stimuli"
trial_list_lst$cat_spa |> 
    gt() |> 
    tab_spanner("Presented word (1)", ends_with("_1")) |> 
    tab_spanner("Presented word (2)", ends_with("_2")) |> 
    cols_label(word_1 = "Orthography",
               ipa_flat_1 = "IPA",
               sampa_1 = "X-SAMPA",
               word_2 = "Orthography",
               ipa_flat_2 = "IPA",
               sampa_2 = "X-SAMPA")
```